### 2016-10-03 Chapters/ResponseTheory.tex
Always link acronyms back to list of acronyms

The derivation above holds for an exact state, a situation that is never
realized in common practice. However, response theory is
also applicable to approximate states.
As noted by \citeauthor{Norman2011-ad}, a response theory treatment of
molecular properties can be summarized into the following four
steps:\autocite{Norman2011-ad}
\begin{enumerate}
    \item Single out a quantity of interest that can be connected to
      molecular properties. In our case, the time-averaged
      quasienergy;
    \item Introduce a suitable parametrization of the time-dependent,
      phase-separated wave function;
    \item Devise the appropriate equation of motion, based on the
      time-dependent Schr\"{o}dinger equation, for the time-dependent
      wave function parameters;
    \item Identify response functions and response equations.
\end{enumerate}
Thus, one can formulate and compute molecular properties based on
the approximate state ans\"{a}tze of quantum chemistry.
An interesting formulation of \acrshort{SCF} response theory is the
open-ended \acrshort{AO} density matrix-based approach presented by
\citeauthor{Thorvaldsen2008-sg}.\autocite{Thorvaldsen2008-sg}
By formulating the perturbation-strenght derivative of the time-averaged
quasienergy Lagrangian in terms of the \acrshort*{AO} density matrix,
the authors were able to formulate arbitrary order response functions
in the \acrshort*{AO} basis.
The density matrix is the variational degree of freedom and the response
equations can be obtained by differentiation of the \gls{TDSCF}
equation, the stationarity condition on the density matrix.
The approach lends itself to an efficient \acrshort*{AO}-based
implementation,\autocite{Larsen2000-hj, Kjaergaard2008-hy} can fully
leverage different response parameters elimination
schemes,\autocite{Thorvaldsen2008-sg, Kristensen2008-hv} and is amenable
to an open-ended, recursive implementation.\autocite{Ringholm2014-gx,
Friese2015-kb, Friese2015-bu}
We will base our derivation of \acrshort*{SCF} response theory for
quantum/classical polarizable Hamiltonians of this framework.

@@

The development of response theory for quantum/classical polarizable
Hamiltonians has moved almost at the same pace as the polarizable
models themselves.
An ample literature exists in the context of the
\acrshort{PCM},\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
polarizable \acrshort{MM} models,\autocite{Curutchet2009-bt, Olsen2010-wa,
Lipparini2012-hx, Lipparini2012-tl}
and general \acrshort*{QM}/\acrshort*{MM}/Continuum models.\autocite{Steindal2011-ki, Caprasecca2012-ir, Lipparini2013-ud}

@@

The variational setting introduced in Section \ref{sec:variational},
offers a series of advantages also in the formulation of response
theory. As already noted, a classical counterpart to the
Hellmann--Feynman theorem holds:
\begin{equation}\tag{\ref{eq:classical-HF} from Chapter \ref{ch:CSM}}
  \deriv{U_\mathrm{pol}}{F} = \pderiv{U_\mathrm{pol}}{F}
  + \pderiv{U_\mathrm{pol}}{\s}\pderiv{\s}{F}
  = \pderiv{U_\mathrm{pol}}{F},
\end{equation}
which forms the basis for our formal development in
\paper{V}.\autocite{pcm-openrsp}
There we show how the variational polarization functional can
successfully achieve a transparent derivation of arbitrary order
response functions, when the fixed-cavity approximation for the
\acrshort*{PCM} is assumed.
Here we report a summary of our findings in \paper{V} employing the
supermatrix form of the classical functional as given in Eq.
\eqref{eq:supermatrix-functional}.
In the notation of \citeauthor{Thorvaldsen2008-sg}, we introduce the
generalized free energy:
\begin{equation}\label{eq:generalized-KS-energy}
  \begin{aligned}
 \tilde{\mathcal{G}}
 &=
 \tilde{\mathcal{E}}
 + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p} +
 {}^t\tilde{\p}\Tr(\tilde{\mathbfup{s}}\pertmat{D}) \\
 &\eqtr
 \left[\pertmat{h} + \pertmat{V}^{t} +
 \frac{1}{2}\pertmat{G}^{\gamma}(\pertmat{D}) -
 \frac{\mathrm{i}}{2}\pertmat{T}\right]\pertmat{D}
 + \tilde{E}_\mathrm{xc}[\tilde{\rho}(\pertmat{D})] + h_\mathrm{nuc}
 + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p}
 + {}^t\tilde{\p}\tilde{\mathbfup{s}}\pertmat{D},
 \end{aligned}
\end{equation}
whose perturbation-strength derivatives at zero perturbation strength
will yield the desired response functions.

To keep track of the number of chain rule applications,
we will denote higher-order derivatives of the \acrshort*{KS}
generalized free energy with the composite ${mn, abc\ldots}$
superscript:
\begin{equation}\label{eq:differential}
  \mathcal{G}^{mn, abc}
  =
  \frac{\partial^{m+n+3}\mathcal{G}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}
  =
    \mathcal{E}^{m, abc} +
  \frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c} \text{.}
\end{equation}
In this notation, the $abc\ldots$ indices denote the perturbations with
respect to which differentiation occurs. The index $m$ denotes the order
of differentiation with respect to the density matrix. $n$
denotes the order of differentiation with respect to the
polarization.
The classical interaction kernel $\V$ might, in the most general
scenario, depend on the molecular geometry.
In \paper{V} and in the following we ignore this possibility and work
under the simplifying assumption that $\V$ is independent of the nuclear
positions.
This goes under the name of \emph{fixed-cavity approximation} in the
context of the \acrshort*{PCM}.\autocite{Cammi1994-qj}
Under this simplifying assumption, the perturbation-strength derivative
of the classical functional will only include terms stemming from the
linear interaction term:
\begin{equation}\label{eq:polarization-derivative}
 \deriv{}{\epsilon_a}\left\lbrace
 U_\mathrm{pol}
 \right\rbrace_T
 \eqavetr
 {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D} \text{,}
\end{equation}
where the second term only involves derivatives of the source integrals.
Moreover, in the general expression for the perturbation-strength
derivative of the polarization functional in Eq. \eqref{eq:differential},
the $m$ and $n$ indices can only assume the values 0 or 1.
In fact, the density matrix dependence is at most linear in the
polarization functional, while the classical Hellmann--Feynman theorem
ensures that the polarization appears at most linearly in $\tilde{\mathcal{G}}^{00,a}$:
\begin{equation}\label{eq:tilde-G00-a}
  \begin{aligned}
    \tilde{\mathcal{G}}^{00,a}(\tilde{\mat{D}}, \tilde{\sigma}, t)
    &= \tilde{\mathcal{E}}^{0,a} +
    {}^t\tilde{\p}\Tr({\tilde{\mathbfup{s}}^a\pertmat{D}}) \\
  &\eqtr
  (\pertmat{h}^a + \pertmat{V}^{t,a} + \frac{1}{2}\pertmat{G}^{\gamma, a}(\pertmat{D}) + \pertmat{F}^{\Omega_a}_\mathrm{xc}
  -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D}
  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}.
  \end{aligned}
\end{equation}
$\pertmat{F}^{\Omega_a}_\mathrm{xc}$ is the functional derivative matrix
defined in terms of the perturbed overlap distributions
$\tilde{\Omega}^a$.
Further perturbation-strength differentiations and evaluation at zero
perturbation strength will yield the desired response functions:
\begin{subequations}
  \begin{align}
    L^a &\eqavetr
    \mathcal{G}^{00,a} - \mat{S}^a\mat{W} \label{eq:first-order} \\
    L^{ab} &\eqavetr \mathcal{G}^{00,ab}
    + \mathcal{G}^{10,a}\mat{D}^{b}
    + \mathcal{G}^{01,a}\p^{b}
    - \mat{S}^{ab}\mat{W}
    - \mat{S}^{a}\mat{W}^{b} \label{eq:second-order} \\
    L^{abc} &\eqavetr
      \mathcal{G}^{00,abc}
    + \mathcal{G}^{10,ac}\mat{D}^{b}
    + \mathcal{G}^{10,ab}\mat{D}^{c}
    + \mathcal{G}^{20,a}\mat{D}^{b}\mat{D}^{c} \nonumber \\
    &+ \mathcal{G}^{10,a}\mat{D}^{bc}
    + \mathcal{G}^{11,a}\mat{D}^{b}\p^c \label{eq:third-order} \\
    &+ \mathcal{G}^{01,ac}\p^{b}
    + \mathcal{G}^{01,ab}\p^{c}
    + \mathcal{G}^{01,a}\p^{bc}
    + \mathcal{G}^{11,a}\p^{b}\mat{D}^c \nonumber \\
    &- \mat{S}^{abc}\mat{W}
    - \mat{S}^{ab}\mat{W}^{c}
    - \mat{S}^{ac}\mat{W}^{b}
    - \mat{S}^{a}\mat{W}^{bc} \nonumber
  \end{align}
\end{subequations}
and similarly for higher orders.

@@

Response equations can be obtained by differentiating the
\acrshort{TDSCF} equation with respect to the perturbation tuple
$b_N$:\autocite{Thorvaldsen2008-sg, Ringholm2014-gx}
\begin{equation}\label{eq:response-equation}
 \left[\genHessian - \omega_{b_N}\genMetric \right]\rspParam{b_N} = \RHS{b_N}
\end{equation}
where the generalized Hessian $\genHessian$ and metric $\genMetric$
matrices were introduced and are defined by their transformations on the
response parameters $\rspParam{b_N}$:\autocite{Larsen2000-hj,
Kjaergaard2008-hy}
\begin{subequations}
\begin{align}
  \genHessian\rspParam{b_N} &=
  \mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})\mat{D}\mat{S}
  -\mat{S}\mat{D}\mat{G}^\mathrm{KS}([\rspParam{b_N},
  \mat{D}]_{\mat{S}}) \nonumber \\
&+\mat{F} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S}
-\mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{F}
\label{eq:E2-lintra} \\
&+ \ASCh{b_N}\mat{\esp}\mat{D}\mat{S}
-\mat{S}\mat{D}\mat{\esp}\ASCh{b_N} \nonumber
\\
  \genMetric\rspParam{b_N} &= \mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S} \text{.}
  \label{eq:S2-lintra}
\end{align}
\end{subequations}
The generalized Hessian matrix $\genHessian$ includes two types of
classical terms: implicit and explicit. \emph{Implicit} terms are those
included in the zeroth-order Fock matrix, $\mat{F}$.
The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
as they explicitly involve the $N$-th order homogenous polarization
variational density.
The nonequilibrium response framework needed in continuum models when
dealing with frequency-dependent properties,\autocite{Cammi1999-rb, Tomasi2005-vm}
will affect the explicit terms in Eq. \ref{eq:E2-lintra} which are
evaluated by making use of the optical permittivity $\dield$ in the
polarization equations \eqref{eq:part-polarization-eq}.
The \acrlong*{RHS} in the response equation only includes terms that
depend on particular contributions up to the desired order or
lower-order perturbed variational densities:\autocite{Ringholm2014-gx}
\begin{equation}\label{eq:RHS}
  \RHS{b_N} =
  \left[\left( \tilde{\mathbfcal{F}}
-
\frac{\mathrm{i}}{2}\pertmat{S}\deriv{}{t}\right)\pertmat{D}\pertmat{S}\right]^{\ominus,
b_N}_\mathrm{P},
\end{equation}
where $\mathbfcal{F}$ is the generalized \acrshort*{KS} matrix, see
\noparcite[ref.][]{Thorvaldsen2008-sg} for details.

>>>>

The derivation above holds for an exact state, a situation that is never
realized in common practice. However, response theory is
also applicable to approximate states.
As noted by \citeauthor{Norman2011-ad}, a response theory treatment of
molecular properties can be summarized into the following four
steps:\autocite{Norman2011-ad}
\begin{enumerate}
    \item Single out a quantity of interest that can be connected to
      molecular properties. In our case, the time-averaged
      quasienergy;
    \item Introduce a suitable parametrization of the time-dependent,
      phase-separated wave function;
    \item Devise the appropriate equation of motion, based on the
      time-dependent Schr\"{o}dinger equation, for the time-dependent
      wave function parameters;
    \item Identify response functions and response equations.
\end{enumerate}
Thus, one can formulate and compute molecular properties based on
the approximate state ans\"{a}tze of quantum chemistry.
An interesting formulation of \acrshort{SCF} response theory is the
open-ended \acrshort{AO} density matrix-based approach presented by
\citeauthor{Thorvaldsen2008-sg}.\autocite{Thorvaldsen2008-sg}
By formulating the perturbation-strenght derivative of the time-averaged
quasienergy Lagrangian in terms of the \acrshort{AO} density matrix,
the authors were able to formulate arbitrary order response functions
in the \acrshort{AO} basis.
The density matrix is the variational degree of freedom and the response
equations can be obtained by differentiation of the \gls{TDSCF}
equation, the stationarity condition on the density matrix.
The approach lends itself to an efficient \acrshort{AO}-based
implementation,\autocite{Larsen2000-hj, Kjaergaard2008-hy} can fully
leverage different response parameters elimination
schemes,\autocite{Thorvaldsen2008-sg, Kristensen2008-hv} and is amenable
to an open-ended, recursive implementation.\autocite{Ringholm2014-gx,
Friese2015-kb, Friese2015-bu}
We will base our derivation of \acrshort{SCF} response theory for
quantum/classical polarizable Hamiltonians of this framework.

@@

The development of response theory for quantum/classical polarizable
Hamiltonians has moved almost at the same pace as the polarizable
models themselves.
An ample literature exists in the context of the
\acrshort{PCM},\autocite{Cammi1994-qj, Cammi1996-wf, Cammi1996-vx,
Cammi1999-rb, Cammi2003-qy, Frediani2005-nc, Ferrighi2010-pm}
polarizable \acrshort{MM} models,\autocite{Curutchet2009-bt, Olsen2010-wa,
Lipparini2012-hx, Lipparini2012-tl}
and general \acrshort{QM}/\acrshort{MM}/Continuum models.\autocite{Steindal2011-ki, Caprasecca2012-ir, Lipparini2013-ud}

@@

The variational setting introduced in Section \ref{sec:variational},
offers a series of advantages also in the formulation of response
theory. As already noted, a classical counterpart to the
Hellmann--Feynman theorem holds:
\begin{equation}\tag{\ref{eq:classical-HF} from Chapter \ref{ch:CSM}}
  \deriv{U_\mathrm{pol}}{F} = \pderiv{U_\mathrm{pol}}{F}
  + \pderiv{U_\mathrm{pol}}{\s}\pderiv{\s}{F}
  = \pderiv{U_\mathrm{pol}}{F},
\end{equation}
which forms the basis for our formal development in
\paper{V}.\autocite{pcm-openrsp}
There we show how the variational polarization functional can
successfully achieve a transparent derivation of arbitrary order
response functions, when the fixed-cavity approximation for the
\acrshort{PCM} is assumed.
Here we report a summary of our findings in \paper{V} employing the
supermatrix form of the classical functional as given in Eq.
\eqref{eq:supermatrix-functional}.
In the notation of \citeauthor{Thorvaldsen2008-sg}, we introduce the
generalized free energy:
\begin{equation}\label{eq:generalized-KS-energy}
  \begin{aligned}
 \tilde{\mathcal{G}}
 &=
 \tilde{\mathcal{E}}
 + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p} +
 {}^t\tilde{\p}\Tr(\tilde{\mathbfup{s}}\pertmat{D}) \\
 &\eqtr
 \left[\pertmat{h} + \pertmat{V}^{t} +
 \frac{1}{2}\pertmat{G}^{\gamma}(\pertmat{D}) -
 \frac{\mathrm{i}}{2}\pertmat{T}\right]\pertmat{D}
 + \tilde{E}_\mathrm{xc}[\tilde{\rho}(\pertmat{D})] + h_\mathrm{nuc}
 + \frac{1}{2}{}^t\tilde{\p}\V\tilde{\p}
 + {}^t\tilde{\p}\tilde{\mathbfup{s}}\pertmat{D},
 \end{aligned}
\end{equation}
whose perturbation-strength derivatives at zero perturbation strength
will yield the desired response functions.

To keep track of the number of chain rule applications,
we will denote higher-order derivatives of the \acrshort{KS}
generalized free energy with the composite ${mn, abc\ldots}$
superscript:
\begin{equation}\label{eq:differential}
  \mathcal{G}^{mn, abc}
  =
  \frac{\partial^{m+n+3}\mathcal{G}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c}
  =
    \mathcal{E}^{m, abc} +
  \frac{\partial^{m+n+3}U_\mathrm{pol}}{\partial(\mat{D}^T)^m\partial\p^n\partial\epsilon_a\partial\epsilon_b\partial\epsilon_c} \text{.}
\end{equation}
In this notation, the $abc\ldots$ indices denote the perturbations with
respect to which differentiation occurs. The index $m$ denotes the order
of differentiation with respect to the density matrix. $n$
denotes the order of differentiation with respect to the
polarization.
The classical interaction kernel $\V$ might, in the most general
scenario, depend on the molecular geometry.
In \paper{V} and in the following we ignore this possibility and work
under the simplifying assumption that $\V$ is independent of the nuclear
positions.
This goes under the name of \emph{fixed-cavity approximation} in the
context of the \acrshort{PCM}.\autocite{Cammi1994-qj}
Under this simplifying assumption, the perturbation-strength derivative
of the classical functional will only include terms stemming from the
linear interaction term:
\begin{equation}\label{eq:polarization-derivative}
 \deriv{}{\epsilon_a}\left\lbrace
 U_\mathrm{pol}
 \right\rbrace_T
 \eqavetr
 {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D} \text{,}
\end{equation}
where the second term only involves derivatives of the source integrals.
Moreover, in the general expression for the perturbation-strength
derivative of the polarization functional in Eq. \eqref{eq:differential},
the $m$ and $n$ indices can only assume the values 0 or 1.
In fact, the density matrix dependence is at most linear in the
polarization functional, while the classical Hellmann--Feynman theorem
ensures that the polarization appears at most linearly in $\tilde{\mathcal{G}}^{00,a}$:
\begin{equation}\label{eq:tilde-G00-a}
  \begin{aligned}
    \tilde{\mathcal{G}}^{00,a}(\tilde{\mat{D}}, \tilde{\sigma}, t)
    &= \tilde{\mathcal{E}}^{0,a} +
    {}^t\tilde{\p}\Tr({\tilde{\mathbfup{s}}^a\pertmat{D}}) \\
  &\eqtr
  (\pertmat{h}^a + \pertmat{V}^{t,a} + \frac{1}{2}\pertmat{G}^{\gamma, a}(\pertmat{D}) + \pertmat{F}^{\Omega_a}_\mathrm{xc}
  -\frac{\mathrm{i}}{2}\pertmat{T}^a)\pertmat{D}
  + h_\mathrm{nuc}^a + {}^t\tilde{\p}\tilde{\mathbfup{s}}^a\pertmat{D}.
  \end{aligned}
\end{equation}
$\pertmat{F}^{\Omega_a}_\mathrm{xc}$ is the functional derivative matrix
defined in terms of the perturbed overlap distributions
$\tilde{\Omega}^a$.
Further perturbation-strength differentiations and evaluation at zero
perturbation strength will yield the desired response functions:
\begin{subequations}
  \begin{align}
    L^a &\eqavetr
    \mathcal{G}^{00,a} - \mat{S}^a\mat{W} \label{eq:first-order} \\
    L^{ab} &\eqavetr \mathcal{G}^{00,ab}
    + \mathcal{G}^{10,a}\mat{D}^{b}
    + \mathcal{G}^{01,a}\p^{b}
    - \mat{S}^{ab}\mat{W}
    - \mat{S}^{a}\mat{W}^{b} \label{eq:second-order} \\
    L^{abc} &\eqavetr
      \mathcal{G}^{00,abc}
    + \mathcal{G}^{10,ac}\mat{D}^{b}
    + \mathcal{G}^{10,ab}\mat{D}^{c}
    + \mathcal{G}^{20,a}\mat{D}^{b}\mat{D}^{c} \nonumber \\
    &+ \mathcal{G}^{10,a}\mat{D}^{bc}
    + \mathcal{G}^{11,a}\mat{D}^{b}\p^c \label{eq:third-order} \\
    &+ \mathcal{G}^{01,ac}\p^{b}
    + \mathcal{G}^{01,ab}\p^{c}
    + \mathcal{G}^{01,a}\p^{bc}
    + \mathcal{G}^{11,a}\p^{b}\mat{D}^c \nonumber \\
    &- \mat{S}^{abc}\mat{W}
    - \mat{S}^{ab}\mat{W}^{c}
    - \mat{S}^{ac}\mat{W}^{b}
    - \mat{S}^{a}\mat{W}^{bc} \nonumber
  \end{align}
\end{subequations}
and similarly for higher orders.

@@

Response equations can be obtained by differentiating the
\acrshort{TDSCF} equation with respect to the perturbation tuple
$b_N$:\autocite{Thorvaldsen2008-sg, Ringholm2014-gx}
\begin{equation}\label{eq:response-equation}
 \left[\genHessian - \omega_{b_N}\genMetric \right]\rspParam{b_N} = \RHS{b_N}
\end{equation}
where the generalized Hessian $\genHessian$ and metric $\genMetric$
matrices were introduced and are defined by their transformations on the
response parameters $\rspParam{b_N}$:\autocite{Larsen2000-hj,
Kjaergaard2008-hy}
\begin{subequations}
\begin{align}
  \genHessian\rspParam{b_N} &=
  \mat{G}^\mathrm{KS}([\rspParam{b_N}, \mat{D}]_{\mat{S}})\mat{D}\mat{S}
  -\mat{S}\mat{D}\mat{G}^\mathrm{KS}([\rspParam{b_N},
  \mat{D}]_{\mat{S}}) \nonumber \\
&+\mat{F} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S}
-\mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{F}
\label{eq:E2-lintra} \\
&+ \ASCh{b_N}\mat{\esp}\mat{D}\mat{S}
-\mat{S}\mat{D}\mat{\esp}\ASCh{b_N} \nonumber
\\
  \genMetric\rspParam{b_N} &= \mat{S} [\rspParam{b_N}, \mat{D}]_{\mat{S}} \mat{S} \text{.}
  \label{eq:S2-lintra}
\end{align}
\end{subequations}
The generalized Hessian matrix $\genHessian$ includes two types of
classical terms: implicit and explicit. \emph{Implicit} terms are those
included in the zeroth-order Fock matrix, $\mat{F}$.
The last two terms in Eq. \ref{eq:E2-lintra} are \emph{explicit} terms,
as they explicitly involve the $N$-th order homogenous polarization
variational density.
The nonequilibrium response framework needed in continuum models when
dealing with frequency-dependent properties,\autocite{Cammi1999-rb, Tomasi2005-vm}
will affect the explicit terms in Eq. \ref{eq:E2-lintra} which are
evaluated by making use of the optical permittivity $\dield$ in the
polarization equations \eqref{eq:part-polarization-eq}.
The \acrlong{RHS} in the response equation only includes terms that
depend on particular contributions up to the desired order or
lower-order perturbed variational densities:\autocite{Ringholm2014-gx}
\begin{equation}\label{eq:RHS}
  \RHS{b_N} =
  \left[\left( \tilde{\mathbfcal{F}}
-
\frac{\mathrm{i}}{2}\pertmat{S}\deriv{}{t}\right)\pertmat{D}\pertmat{S}\right]^{\ominus,
b_N}_\mathrm{P},
\end{equation}
where $\mathbfcal{F}$ is the generalized \acrshort{KS} matrix, see
\noparcite[ref.][]{Thorvaldsen2008-sg} for details.
